### 2025-11-10 revision-2/editor_letter.tex
GK corrections

\documentclass[a4paper,10pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{lipsum}
\usepackage{xcolor}
\def\red#1{\textcolor{black}{#1}}

\begin{document}

% Sender Information
% \begin{flushright}
%     \begin{tabular}{l}
%         \textbf{Piotr Stasiak} \\
%         School of Mathematics, Statistics and Physics\\
%         Newcastle University \\
%         Newcastle  upon Tyne \\
%         p.stasiak@newcastle.ac.uk \\
%         \today
%     \end{tabular}
% \end{flushright}

% \vspace{1cm}

% Recipient Information
The Editors,\\
Physical Review Letters\\
APS\\

\vspace{1cm}

Dear Editor,

\vspace{0.5cm}

We would like first to acknowledge the Refeerees for reading our last submission. We would like to request a third round of referral of our manuscript. Our request is based on the following points.

-  Referee A, has showed themself very positive about our work. In their first report, the Referee says ``\emph{While this finding is intriguing, several issues must be addressed before a recommendation for publication can be made.}'' Whereas in their second review, they concluded their report with ``\emph{This stage is premature to recommend to Physical Review Letters}'', which is not a rejection per se, but rather a request for for further improvements.  In their second report, there are only minor issues left, that we have carefully addressed in our answers. Moreover, we have included in this revised version of the manuscript an analysis of about fifty new reconnection events having a different initial topology. The new set of simulations confirms our original results and makes our results extremely robust.

- Referee B has shown some  bias and a non-professional behavior from their first report. The Referee has tried to build a complete case about one single sentence on our conclusions, which was indented to open a possible question leading to a new research project. We have tried to explain this issue to the Referee in our first answer, but they seem to have completely ignore it, maintaining their unprofessional aggressivity. Furthermore, it seems to us, that the Referee has tried to bring new arguments, simply in order to push down the manuscript, confirming our belief that the Referee has some type of conflict of interest from the beginning. For the sake of professionalism, we have written an answer to Referee B last report, but we are reluctant to think that their bias could change.

We therefore think that referral process has not been fair, and we would like thus to request that our revised version is sent back to Referee A and third Referee is solicited.

Thank you for considering this manuscript; we look forward to your response.

\vspace{1cm}

\textbf{Piotr Stasiak}\\
\textbf{Andrew Baggaley}\\
\textbf{Carlo Barenghi}\\
\textbf{Giorgio Krstulovic}\\
\textbf{Luca Galantucci}\\
\end{document}