### 2025-11-17 revision-2/editor_letter.tex

\documentclass[a4paper,10pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{lipsum}
\usepackage{xcolor}
\def\red#1{\textcolor{black}{#1}}

\begin{document}

% Sender Information
% \begin{flushright}
%     \begin{tabular}{l}
%         \textbf{Piotr Stasiak} \\
%         School of Mathematics, Statistics and Physics\\
%         Newcastle University \\
%         Newcastle  upon Tyne \\
%         p.stasiak@newcastle.ac.uk \\
%         \today
%     \end{tabular}
% \end{flushright}

% \vspace{1cm}

% Recipient Information
The Editors,\\
Physical Review Letters\\
APS\\

\vspace{1cm}

Dear Editor,

\vspace{0.5cm}

We would like first to thank the Refeerees for reading our last submission. We would like to request a third round of referral of our manuscript. Our request is based on the following points.

-  Referee A, has been very positive about our work. In the first report, 
Referee A said ``\emph{While this finding is intriguing, several issues 
must be addressed before a recommendation for publication can be made.}'' 
In the second review, Referee A concluded with 
``\emph{This stage is premature to recommend to Physical Review Letters}'', 
which is not a rejection per se, but rather a request for for further 
improvements.  In their second report, there are only minor issues left, 
that we have now carefully addressed in our answers. Moreover, 
in this newly revised version of the manuscript we have incuded an 
analysis of about fifty new reconnection events having a different 
initial topology. The new set of simulations confirms our original 
results and makes our results extremely robust.

- Referee B has shown bias and non-professional behavior from their first 
report. The Referee has tried to build a complete case against our work
based on a single sentence in our conclusion section, which was only indented 
to open a possible question leading to a new research project. 
We explained this issue in our first revision, but the the Referee
unprofessionally ignored it. Furthermore, it seems to us that the Referee 
has tried to bring new arguments simply in order to push down the manuscript, 
confirming our belief that there is some conflict of interest.
For professional sake, we answered the last report of Referee B, 
but we doubt that their bias could change.

We therefore think that referral process has not been entirely fair: 
we would like 
thus to request that our revised version is sent back to Referee A and 
third Referee is solicited.

Thank you for considering this manuscript; we look forward to your response.

\vspace{1cm}

\textbf{Piotr Stasiak}\\
\textbf{Andrew Baggaley}\\
\textbf{Carlo Barenghi}\\
\textbf{Giorgio Krstulovic}\\
\textbf{Luca Galantucci}\\
\end{document}